---
title: "House price"
author: "ForozanHeidaryan"
date: "2022-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This data set is about the sale price of buildings in the city of Ame Iva which has been 
Collected by statistics named Dean de cock 

# Understanding the Business Question

Price Recommendation for  house
Data inspection

# Dataset Description
## File descriptions


data_description.txt - full description of each column, originally prepared by Dean De Cock

but lightly edited to match the column names used here

sample_submission.csv - a benchmark submission from a linear regression on year and month of 
sale, lot square footage, and number of bedrooms

Data fields

Here's a brief version of what you'll find in the data description file.

SalePrice :   the property's sale price in dollars. This is the target variable that you're 

trying to predict.

MSSubClass:   The building class

MSZoning:     The general zoning classification

LotFrontage:  Linear feet of street connected to property

LotArea:      Lot size in square feet

Street:       Type of road access

Alley:        Type of alley access

LotShape:     General shape of property

LandContour:  Flatness of the property

Utilities:    Type of utilities available

LotConfig:    Lot configuration

LandSlope:    Slope of property

Neighborhood: Physical locations within Ames city limits

Condition1:   Proximity to main road or railroad

Condition2:   Proximity to main road or railroad (if a second is present)

BldgType:     Type of dwelling

HouseStyle:   Style of dwelling

OverallQual:  Overall material and finish quality

OverallCond:  Overall condition rating

YearBuilt:    Original construction date

YearRemodAdd: Remodel date

RoofStyle:    Type of roof

RoofMatl:     Roof material
Exterior1st:  Exterior covering on house

Exterior2nd:  Exterior covering on house (if more than one material)

MasVnrType:   Masonry veneer type

MasVnrArea:   Masonry veneer area in square feet

ExterQual:    Exterior material quality

ExterCond:    Present condition of the material on the exterior

Foundation:   Type of foundation

BsmtQual:     Height of the basement

BsmtCond:     General condition of the basement

BsmtExposure: Walkout or garden level basement walls

BsmtFinType1: Quality of basement finished area

BsmtFinSF1:   Type 1 finished square feet

BsmtFinType2: Quality of second finished area (if present)

BsmtFinSF2:   Type 2 finished square feet

BsmtUnfSF:    Unfinished square feet of basement area

TotalBsmtSF:  Total square feet of basement area

Heating:      Type of heating

HeatingQC:    Heating quality and condition

CentralAir:   Central air conditioning

Electrical:   Electrical system

1stFlrSF:     First Floor square feet

2ndFlrSF:     Second floor square feet

LowQualFinSF: Low quality finished square feet (all floors)

GrLivArea:    Above grade (ground) living area square feet

BsmtFullBath: Basement full bathrooms

BsmtHalfBath: Basement half bathrooms

FullBath:     Full bathrooms above grade

HalfBath:     Half baths above grade

Bedroom:      Number of bedrooms above basement level

Kitchen:      Number of kitchens

KitchenQual:  Kitchen quality

TotRmsAbvGrd: Total rooms above grade (does not include bathrooms)

Functional:   Home functionality rating

Fireplaces:   Number of fireplaces

FireplaceQu:  Fireplace quality

GarageType:   Garage location

GarageYrBlt:  Year garage was built

GarageFinish: Interior finish of the garage

GarageCars:   Size of garage in car capacity

GarageArea:   Size of garage in square feet

GarageQual:   Garage quality

GarageCond:   Garage condition

PavedDrive:   Paved driveway

WoodDeckSF:   Wood deck area in square feet

OpenPorchSF:  Open porch area in square feet

EnclosedPorch:Enclosed porch area in square feet

3SsnPorch:    Three season porch area in square feet

ScreenPorch:  Screen porch area in square feet

PoolArea:     Pool area in square feet

PoolQC:       Pool quality

Fence:        Fence quality

MiscFeature:  Miscellaneous feature not covered in other categories

MiscVal:      $Value of miscellaneous feature

MoSold:       Month Sold

YrSold:       Year Sold

SaleType:     Type of sale

SaleCondition:Condition of sale

### read data from file 
```{r}
data <- read.csv("train.csv" , header = TRUE )
```

summary of the data

```{r}
dim(data)
length(unique(data$Id))
summary(data)
```

Convert categorical variables to factor

```{r}
cat_var <- c("MSZoning" , "Street" , "Alley" ,"LotShape" , "LandContour", "Utilities" , "LotConfig" , "LandSlope" ,
             "Neighborhood" , "Condition1" , "Condition2" , "BldgType" ,
             "HouseStyle", "OverallQual", "OverallCond" , "RoofStyle" , "Exterior1st" , "Exterior2nd" , "MasVnrType",
             "ExterQual" ,"ExterCond" , "Foundation" , "BsmtQual" ,"BsmtCond" ,
             "BsmtExposure" ,"BsmtFinType1" ,"BsmtFinType2" , "Heating" , "HeatingQC" , "CentralAir" , "Electrical" , 
             "KitchenQual" , "Functional" , "Fireplaces" , "FireplaceQu" ,
             "GarageType" , "GarageFinish" , "GarageCars" , "GarageQual" ,"GarageCond" , "PavedDrive" , "PoolQC", 
             "Fence","RoofMatl" , 
             "MiscFeature" , "SaleType" , "SaleCondition")
```

```{r}
data[,cat_var]  <-  lapply(data[,cat_var] , factor)
knitr::kable(summary(data)) 
knitr::kable(str(data))
```

Identification missing values

```{r}
mv_summary2        <-  data.frame('variables name' = colnames(data))
mv_summary2$freq   <- apply(data , 2 , function(x)  sum(is.na(x)))
mv_summary2$pers   <- round(mv_summary2$freq / nrow(data) , 3) * 100
mv_summary2_1      <- as.data.frame(mv_summary2[mv_summary2$pers > 0 & mv_summary2$pers <10, ])
mv_summary2_2      <- as.data.frame(mv_summary2[mv_summary2$pers > 10 , ])
knitr::kable(mv_summary2_1)
knitr::kable(mv_summary2_2)
```

Removing columns that have more than 10% missing value
In my opinion, these columns have high missing values and cause problems in over modelling

```{r}
t1 <- data[,c("LotFrontage","Alley","FireplaceQu" ,"Fence" , 
                "MiscFeature" , "PoolQC")]
```


```{r}
data1 <- data[,-which(data %in% t1)]
dim(data1)
```

I also removed all rows that contained missing values. this is the  easiest category to get rid of missing values

```{r}
data2 <- data1[apply(data1,1,function(x) any(is.na(x))) == F,]

```

For convenience, I divided the data into continuous and discrete parts

```{r}
cat <- data2[,c("MSZoning" , "Street" ,"LotShape" , "LandContour", "Utilities" , "LotConfig" , "LandSlope" ,
                 "Neighborhood" , "Condition1" , "Condition2" , "BldgType" ,
                 "HouseStyle", "OverallQual", "OverallCond" , "RoofStyle" , "Exterior1st" , "Exterior2nd" , 
                 "MasVnrType"    ,
                 "ExterQual" ,"ExterCond" , "Foundation" , "BsmtQual" ,"BsmtCond" ,
                 "BsmtExposure" ,"BsmtFinType1" ,"BsmtFinType2" , "Heating" , "HeatingQC" , "CentralAir" ,
                 "Electrical"  , 
                 "KitchenQual" , "Functional" , "GarageType" , "GarageFinish" , "GarageCars" , "GarageQual" ,
                 "GarageCond" , "PavedDrive" , 
                 "RoofMatl" , 
                 "SaleType" , "SaleCondition" ,"Fireplaces" )]
```

I prefer to convert the columns that contain the date field to age, this helps me more easily determine the relationship between the price and the life of the house.

```{r}

data3 <- data2[,-c(18,19,57,72)]
today <- as.Date("2022"  , format = "%Y")
```

```{r}
data3$ageBuilt <- as.Date(as.character(data2$YearBuilt) , format = "%Y")

data3$ageBuilt   <- as.numeric(today - data3$ageBuilt)

data3$ageBuilt<- round(data3$ageBuilt/365 )
summary(data$ageBuilt)

data3$ageRemodAdd  <- as.Date(as.character(data2$YearRemodAdd), format = "%Y")

data3$ageRemodAdd <- as.numeric(today - data3$ageRemodAdd)

data3$ageRemodAdd <- round(data3$ageRemodAdd / 365 )

summary(data3$ageRemodAdd)

data3$GarageageBlt   <- as.Date(as.character(data2$GarageYrBlt),format = "%Y")
data3$GarageageBlt   <- as.numeric(today - data3$GarageageBlt)
data3$GarageageBlt   <-round(data3$GarageageBlt / 365 )

summary(data3$GarageageBlt)
#Regarding the sale date, it also helps me to better recognize the rise and fall of prices
data3$ageSold  <- as.Date(as.character(data2$YrSold) , format = "%Y")
data3$ageSold  <- as.numeric(today - data3$ageSold)
data3$ageSold  <- round(data3$ageSold / 365 )
summary(data3$ageSold)
```

Examining the distribution of the response variable

```{r}
par(mfrow = c(1,1))
hist(data$SalePrice , breaks = 50,probability = TRUE , col = "#4B0082")
lines(density(data$SalePrice) , col = "red")
```

```{r}
qqnorm(data$SalePrice , col = "#4B0082")
qqline(data$SalePrice , col = "red")
```

```{r}
par(mfrow = c(1,1))
boxplot(data$SalePrice)
```

```{r}
library("moments")

jarque.test(data$SalePrice)
#pvalue < 0   -> h0 regect
```

Graphs and statical tests indicate  that the response variable does not  follow 
A normal distribution  which is quite  reasonable for house prices


I would like to make the response variable as close to a normal distribution 
As possible this may help me in the modelling  for this I use the box cox transformation
```{r}

library("MASS")
box_result <- boxcox(data$SalePrice ~ 1 , lambda = seq(-5 , 5 , 0.1))
box_result <- data.frame(box_result)
lambda <- box_result[which(box_result$y == max(box_result$y)),] 
#It observes that zero is inside the confidence interval, so I use logarithm variable change 
data3$SalePrice <- log(data3$SalePrice)
```
It is clear that zero is inside this confidence interval
Therefore, I use logarithmic transformation

```{r}
hist(data3$SalePrice , breaks = 50 , probability = TRUE , col = "#4B0082")
lines(density(data3$SalePrice) , col = "red")
```

```{r}
qqnorm(data3$SalePrice , col = "#4B0082")
qqline(data3$SalePrice , col = "red")
```

```{r}
jarque.test(data3$SalePrice)
```

```{r}
anscombe.test(data3$SalePrice)
```

```{r}
par(mfrow = c(1,1))
boxplot(data3$SalePrice ,main = "plot respont", col = "#4B0082")
```
It seems that the data is far from the normal distribution

```{r}
tukey_u <- quantile(data3$SalePrice , probs = 0.75) + 1.5 * IQR(data3$SalePrice)

sum(data3$SalePrice > tukey_u)
```
For convenience, I divided the data into continuous and discrete parts

```{r}
count <- data3[,-which(data3 %in% cat)]
cunt1 <- count[,c(1:17)]
cunt2 <- count[,c(18:29)]

cunt1 <- cunt1[,-1]
cunt1$SalePrice  <- cunt2$SalePrice
```

Histograms relate to continuous variables
```{r}
par(mar = c(2,2,2,2))
par(mfrow = c(4,4))
for (i in 1:16 ) {
  hist(cunt1[,i] , xlab =   "" ,col = "#4B0082", main = paste("HIST COUNT 1" ,
                                                              names(cunt1[i])))
}
```

```{r}
par(mar = c(2,2,2,2))
par(mfrow = c(4,4))
for (i in 1:11) {
  hist(cunt2[,i] , xlab = "" , col = "#4B0082" , main = paste("HIST COUNT 2" , 
                                                              names(cunt1[i])))
}
```

Correlation of continuous variables versus response variable  

```{r}
cros_tab <- round(cor(count),2)

library("corrplot")
par(mfrow = c(1,1))
corrplot(cros_tab)
```

```{r}
par(mar=c(2,2,2,2))
par(mfrow = c(4,4))
for (i in 1:16) {
  plot(cunt1[,i] , cunt2$SalePrice,xlab = "" , main = paste("plot" ,names(cunt1)[i] ))
}
```

```{r}
par(mar=c(2,2,2,2))
par(mfrow = c(4,4))
for (i in 1:11) {
  plot(cunt2[,i] , cunt2$SalePrice,xlab = "" , main = paste("plot" ,names(cunt2)[i] ))
}
```

We go  to the distribution of discrete variables

```{r}
categori <- data3[,which( data3 %in% cat)]
colnames(categori)
dim(categori)
```

Examining the distribution of discrete variables


It is observed that  residential low-density and residential following villages have the highest average price, of course, it is obvious that these two types have a higher frequency than the other 
```{r}
table(data3$MSZoning)
boxplot(SalePrice ~ MSZoning , data = data3 , col = "#4B0082" , border = "#7E3817")

```
   
   It is observed that  residential low-density and residential following villages have the highest average price, of course, it is obvious that these two types have a higher frequency than the other 
   
```{r}
table(data3$Street)
boxplot(SalePrice ~  Street , data = data3 , col = "#4B0082" , border = "#7E3817")
```
Buildings that are  located on paved streets have a higher 	average price,although according to the table it is evident that they are mor expensive than other type, so no decision  can be made about this

```{r}
table(data3$LotShape)
boxplot(SalePrice ~ LotShape , data = data3 , col = "#4B0082" , border = "#7E3817")
```

if the shape of the building more regular, it will have higher average 

```{r}
table(data3$LandContour)
boxplot(SalePrice ~ LandContour , data = data3 , col = "#4B0082" , border = "#7E3817") 
```

buildings that are higer than the grouand level ( compared to the street  ) they supposed  to have higher average price 

```{r}
table(data3$Utilities)
boxplot(SalePrice ~  Utilities , data = data3 , col = "#4B0082" , border = "#7E3817")
```

I cant make any decisions  about this because  I only have one sample of the nosewat type _ I’ll probablity drop this variable altogether because it cant be a good explanation  

```{r}
table(data3$LotConfig)
boxplot(SalePrice ~  LotConfig , data = data3 , col = "#4B0082" , border = "#7E3817")
```

this variable vague for me, may be I will  catch some thing in the moddeling process. Ather wise I’ll delete irt

```{r}
table(data3$LandSlope)
boxplot(SalePrice ~  LandSlope , data = data3 , col = "#4B0082" , border = "#7E3817")
```

 the avrage of the variable is close to each other, and it seems that the presence of outlier variable in the gentle slope is due to the high frequency of this type, according  to observing image of Ames, we found that this city is flat - in the case of colinearity I will delete it 

```{r}
table(data3$Neighborhood)
boxplot(SalePrice ~ Neighborhood , data = data3 , col = "#4B0082" , border = "#7E3817")
```

```{r}
table(data3$Condition1)
boxplot(SalePrice ~ Condition1 , data = data3 , col = "#4B0082" , border = "#7E3817")
```
most of the building are in normal location. In the case of(arteria street) I suspect that the noise of the cars is involved in lowering the price 

```{r}
table(data3$Condition2)
boxplot(SalePrice ~ Condition2 , data = data3 , col = "#4B0082" , border = "#7E3817")
```
 conditions1 and 2 are very similar and I have the same opinion about these two conditions

```{r}
table(data3$BldgType)
boxplot(SalePrice ~ BldgType , data = data3 , col = "#4B0082" , border = "#7E3817")
```
single_family detached house and town houses inside units have a higher average price than others.and the houses that have been converted for these two families_ have the 
Lowest  price

```{r}
table(data3$HouseStyle)
boxplot(SalePrice ~ HouseStyle , data = data3 , col = "#4B0082" , border = "#7E3817")
```
house style raises all issues related to the high square footage and floors of the average house price 
```{r}
table(data3$OverallQual)
boxplot(SalePrice ~ OverallQual , data = data3 , col = "#4B0082" , border = "#7E3817")
```
the overall quality of the building raises the average price and standard deviation 

```{r}
table(data3$OverallCond)
boxplot(SalePrice ~ OverallCond , data = data3 , col = "#4B0082" , border = "#7E3817")
```
of course the better the overall condition of the house, the higher the average price should be, but what is evident here is that houses with the average condition have a higher average price, maybe it is because of the abundance of this group
```{r}
table(data3$RoofStyle)
boxplot(SalePrice ~ RoofStyle , data = data3 , col = "#4B0082"  , border = "#7E3817")
```
 the average is close to each other 

```{r}
table(data3$RoofMatl)
boxplot(SalePrice ~  RoofMatl , data = data3 , col = "#4B0082" , border = "#7E3817")
```
almost all types of standard (composite) single has been formed. I don’t this variable is a good explanation think
```{r}
table(data3$Exterior1st)
boxplot(SalePrice ~ Exterior1st , data = data3 , col = "#4B0082" , border = "#7E3817")
```
Houses covered with  stone and concrete have a relatively higher average price than house covered with wood 

```{r}
table(data3$Exterior2nd)
boxplot(SalePrice ~ Exterior2nd , data = data3 , col = "#4B0082" , border = "#7E3817")
```
Vinyl siding covers most of the house, the rest of the covers are around this average 
```{r}
table(data3$MasVnrType)
boxplot(SalePrice ~ MasVnrType , data = data3 , col = "#4B0082" , border = "#7E3817")
```
stone cladding has the highest average price and brick common has the lowest price 

```{r}
table(data3$ExterQual)
boxplot(SalePrice ~ ExterQual ,data = data3 , col = "#4B0082" , border = "#7E3817")
```
the excellent quality of the material, for the exterior of the building, has the highest average price, and the lower the quality of the materials, the lower the average price

```{r}
table(data3$ExterCond)
boxplot(SalePrice ~ ExterCond , data = data3 , col = "#4B0082" , border = "#7E3817")
```
The variabels distribution is almost similar to the extent qual 
```{r}
table(data3$Foundation)
boxplot(SalePrice ~ Foundation , data = data3 , col = "#4B0082" , border = "#7E3817") 
```
poured concrete and finder block from the most frequent and also have the highest average price

```{r}
table(data3$BsmtQual)
boxplot(SalePrice ~ BsmtQual , data = data3 , col = "#4B0082" , border = "#7E3817")
```
 what so ever the height of the basement is high, the average price rise
```{r}
table(data3$BsmtCond)
boxplot(SalePrice ~ BsmtCond , data = data3 , col = "#4B0082" , border = "#7E3817")
```
the overall quality of the basement raises the average price
```{r}
table(data3$BsmtExposure)
boxplot(SalePrice ~ BsmtExposure , data = data3 , col = "#4B0082" , border = "#7E3817")
```
the quality of the pavement be better the average price will raise

```{r}
table(data3$BsmtFinType1)
boxplot(SalePrice ~ BsmtFinType1 , data = data3 , col = "#4B0082" , border = "#7E3817")
```
As the area of the land increases, the average price increases, of course 
It can be in line with the total area of the building 

```{r}
table(data3$BsmtFinType2)
boxplot(SalePrice ~ BsmtFinType2 , data = data3 , col = "#4B0082" , border = "#7E3817")
```
 the quality of the land has a direct relationship  with the price, but apparently it has formed the majority of the land (probability of multi-collinearity and lack of explicit explanation  )
 
```{r}
table(data3$Heating)
boxplot(SalePrice ~ Heating ,data = data3 , col = "#4B0082" , border = "#7E3817")
```
Houses with gas have the most abundance and the highest price 

```{r}
table(data3$HeatingQC)
boxplot(SalePrice ~ HeatingQC , data = data3 , col = "#4B0082" , border = "#7E3817")
```
the quality of heating has a direct relationship with the price 

```{r}
table(data3$CentralAir)
boxplot(SalePrice ~ CentralAir ,data = data3 , col = "#4B0082" , border = "#7E3817")

```
Having central air conditioning has a direct relationship with the average price 

```{r}
table(data3$Electrical)
boxplot(SalePrice ~ Electrical , data = data3 , col = "#4B0082" , border = "#7E3817")
```
 Most houses with SBrkr have a higher average price 

```{r}
table(data3$KitchenQual)
boxplot(SalePrice ~ KitchenQual , data = data3, col = "#4B0082" , border = "#7E3817")
```
the quality of kitchens has a direct relationship with the price

```{r}
table(data3$Functional)
boxplot(SalePrice ~ Functional, data = data3 , col = "#4B0082" , border = "#7E3817")
```
this category is vague to me maybe I didn’t comment on it and maybe I catch sight in the modelling process
```{r}
table(data3$Fireplaces)
boxplot(SalePrice ~ Fireplaces , data = data3 , col = "#4B0082" , border = "#7E3817")
```
the number of fireplaces has a direct relationship with the price of the house maybe it is in line with the square footage 
```{r}
table(data3$GarageType)
boxplot(SalePrice ~ GarageType , data = data3 , col = "#4B0082" , border = "#7E3817")
```
Garages inside the house and attached to the house have a higher price and houses with 
A separate garage and without a garage have a lower average price 
```{r}
table(data3$GarageFinish)
boxplot(SalePrice ~ GarageFinish , data = data3 , col = "#4B0082" , border = "#7E3817")
```
 complete garages have a higher average price than unfinished garages

```{r}
table(data3$GarageCars)
boxplot(SalePrice ~ GarageCars , data = data3 , col = "#4B0082" , border = "#7E3817")
```
the number of garages has a direct relationship with the price but in the case of houses with 4 garages, a price drop is observed
```{r}
table(data3$GarageCond)
boxplot(SalePrice ~ GarageCond , data = data3 , col = "#4B0082" , border = "#7E3817")
```
paved and equipped sidewalks have a direct relationship with the price 

```{r}
table(data3$PavedDrive)
boxplot(SalePrice ~ PavedDrive , data = data3 , col = "#4B0082" , border = "#7E3817")
```
Normal and conditional sales have the most frequent and average prices and are higher in the sale type, but sales with advance payment and also newly built houses have a higher price 
```{r}
table(data3$SaleType)
boxplot(SalePrice ~ SaleType , data = data3 , col = "#4B0082" , border = "#7E3817")
```
the average of all prices are close to each other, except partial, which may be due to home renovation
```{r}
table(data3$SaleCondition)
boxplot(SalePrice ~ SaleCondition , data = data3 , col = "#4B0082" , border = "#7E3817")
```

Go to modelling 

